---
title: "p8105_hw2_ys3765"
author: "Yixiao Sun"
date: "2023-09-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r hide=TRUE}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)
```

#Problem 2

```{r clean mr_trash_wheel}
mr_trash_wheel <- read_excel("202309 Trash Wheel Collection Data.xlsx",sheet = 1) %>%                       janitor::clean_names() %>% select(-x15,-x16) %>% 
                  mutate(year = as.character(year)) %>%                                                     mutate(month = as.character(month))

mr_trash_wheel<- na.omit(mr_trash_wheel)
mr_trash_wheel$homes_powered <- mr_trash_wheel$weight_tons*500/30
mr_trash_wheel
```

```{r clean prof_trash_wheel}
prof_trash_wheel <- read_excel("202309 Trash Wheel Collection Data.xlsx",sheet = 2) %>%
                    janitor::clean_names()%>%
                    mutate(year = as.character(year)) %>% 
                    mutate(month = as.character(month))
prof_trash_wheel<-prof_trash_wheel[-107,]
prof_trash_wheel$homes_powered <- prof_trash_wheel$weight_tons*500/30
prof_trash_wheel
```


```{r clean gwynnda}
gwynnda <- read_excel("202309 Trash Wheel Collection Data.xlsx",sheet = 4) %>%
          janitor::clean_names()%>%mutate(year = as.character(year)) %>% 
          mutate(month = as.character(month))

gwynnda<-gwynnda[-107,]
gwynnda<-gwynnda[-107,]
gwynnda$homes_powered <- gwynnda$weight_tons*500/30
gwynnda
```

```{r}
mr_trash_wheel$trash_wheel <- "Mr. Trash Wheel"
prof_trash_wheel$trash_wheel <- "Professor Trash Wheel"
gwynnda$trash_wheel <- "Gwynnda"
```

```{r}
combined_data_step1 <- bind_rows(mr_trash_wheel, prof_trash_wheel)
combined_data<-bind_rows(combined_data_step1,gwynnda)
combined_data
```

For these three dataset, the mr_trash_wheel dataset has `r nrow(mr_trash_wheel)` rows and `r ncol(mr_trash_wheel)` after cleaning. The prof_trash_wheel dataset has `r nrow(prof_trash_wheel)` rows and `r ncol(prof_trash_wheel)` after cleaning. The Gwynnda dataset has `r nrow(gwynnda)` rows and `r ncol(gwynnda)` after cleaning. The combined dataset has intotal `r nrow(combined_data)` and `r ncol(combined_data)`. The important variables in each of these datasets including the geographic info of the dumpster, but in order to calculate the trash that could be using for powering homes, weight_tons variable is the most essential one. The total weight of trash collected by Professor Trash Wheel dataset is `r sum(prof_trash_wheel$weight_tons)`. And for the total number of cigarette butts collected by Gwynnda in July of 2021, the number is `r sum(head(gwynnda$cigarette_butts,5))`.




#Problem 3
```{r import data}

library(readr)
mci_baseline_original <- read_csv("data_mci/MCI_baseline.csv",skip=1)%>%
                janitor::clean_names()

mci_baseline <- read_csv("data_mci/MCI_baseline.csv",skip=1)%>%
                janitor::clean_names() %>% 
                mutate(sex = ifelse(sex == 1,"Male","Female")) %>% 
                mutate(apoe4 = ifelse(apoe4 == 1, "Carrier","Non-Carrier")) %>%        
                filter(current_age < age_at_onset|age_at_onset==".")%>%
                mutate(age_at_onset = if_else(age_at_onset == ".", NA, age_at_onset))

mci_baseline

mci_amyloid <- read_csv("data_mci/mci_amyloid.csv",skip = 1)%>%
              janitor::clean_names()%>%
              rename("id" = "study_id")%>%
              pivot_longer(baseline:time_8,
                           names_to = "year",
                           values_to = "value" )%>%
              mutate(year =  case_match(
                              year,
                            "baseline" ~ 0, 
                            "time_2" ~ 2, 
                            "time_4" ~ 4,
                            "time_6" ~ 6, 
                            "time_8" ~ 8))

mci_amyloid

only_data<-anti_join(mci_baseline,mci_amyloid)
only_data_2<-anti_join(mci_amyloid,mci_baseline)
only_data
only_data_2

Combined_data_2<-inner_join(mci_baseline,mci_amyloid)
Combined_data_2

write_csv(Combined_data_2,"/Users/elliot-sun/Desktop/P8105_ Data Science/p8105_hw2_ys3765/data_mci/combined_data.csv")
```
For the cleaning process, I first change the the second dataset's `study_id` into `id` in order to better combine the two datasets, then I change the `sex` and `apoe4` columns in the first dataset from character into numeric. Then to better clean the amyloid dataset, I chose pivot_longer function and renamed the columns of the dataset to utlize better with all the information. Anti_join function was used for finding the number of people who only appeared once in each dataset. And finally I used inner_join function to fully combined these two datasets.

For these two datasets, the first cleaning step is to find those patients who do not show MCI at baseline. The baseline dataset showing features of the patients' geographical info including educational level, age, the onset age. So the basic cleaning step should be using the current age to compare with the age at onset, and we get `r nrow(mci_baseline)` rows of the dataset, comparing with the original dataset has `r nrow(mci_baseline_original) ` numbers of participants recrutied. The average baseline age is `r mean(mci_baseline$current_age)`. The proportion of women in the study are APOE4 carriers are `r sum(mci_baseline$sex == "Female" & mci_baseline$apoe4 == "Carrier")/sum(mci_baseline$sex == "Female")`.

For the two datasets, the mci_baseline has `r nrow(mci_baseline)` rows before cleaning, and for the mci_amyloid has `r nrow(mci_amyloid)` rows. So partial participants appeared in only one dataset.(`r nrow(only_data)` and `r nrow(only_data_2)` from the function only_data and only_data_2). The combined dataset cleaned the data and put the same patients' value together in the order of time length(from beginning to 8 years), the whole dataset contains `r nrow(Combined_data_2)` all other demographic info are shown alone with the age and ID and values.